\documentclass[11pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{latexsym,amsbsy,amssymb,amsmath,amsfonts}
\usepackage{amsthm}
\usepackage{epsfig}
\usepackage{euscript}
\usepackage{fullpage}
\usepackage{multirow,multicol,booktabs}
\usepackage[bf]{caption}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{url}
\usepackage{hyperref}
\hypersetup{
 pdfauthor={Christian Reitwießner},
 pdftitle={MapReduce for plasma-Blockchains}
  unicode,
  breaklinks,
  colorlinks=false,
  pdfborder={0 0 0}
}

\date{}

\renewcommand{\captionfont}{\small}



\hfuzz=0mm
\tolerance=10000
\hbadness=1000


\setlength{\parindent}{0mm}
\setlength{\parskip}{2ex plus0.5ex minus0.5ex}


\newcommand{\Sum}{\sum\limits}
\newcommand{\Prod}{\prod\limits}


\newtheorem{dummytheorem}{Dummy-Theorem}[section]
\newtheorem{definition}[dummytheorem]{Definition}
\newtheorem{lemma}[dummytheorem]{Lemma}
\newtheorem{theorem}[dummytheorem]{Theorem}
\newtheorem{proposition}[dummytheorem]{Proposition}
\newtheorem{property}[dummytheorem]{Property}
\newtheorem{corollary}[dummytheorem]{Corollary}
\newtheorem{example}[dummytheorem]{Example}
\newtheorem{remark}[dummytheorem]{Remark}
\newtheorem{fact}[dummytheorem]{Fact}
\newtheorem{claim}[dummytheorem]{Claim}
\newtheorem{subclaim}{Subclaim}[dummytheorem]
\newtheorem{conjecture}[dummytheorem]{Conjecture}

\newcommand{\uint}{\mathbb{N}}
\newcommand{\rational}{\mathbb{Q}}

\newcommand{\oli}[1]{\overline{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\newcommand{\eps}{\varepsilon}
\newcommand{\card}[1]{\##1}
\newcommand{\length}[1]{\mathrm{length}({#1})}

\newcommand{\pn}[1]{\textnormal{#1}}

\newcommand{\norm}[1]{||{#1}||}
\newcommand{\Norm}[1]{\left|\left|{#1}\right|\right|}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{document}
\selectlanguage{english}

\title{Computation on the Edges of Plasma-Blockchains}

\author{Christian Reitwießner\\
{\tt chris@ethereum.org}}


\maketitle

\begin{abstract}
We propose a programming paradigm to be used for smart contract blockchain
system arranged in a tree-like structure as in the plasma~\cite{plasma} system.
We show how smart contracts can be written so that invariants global to the
whole system are kept, even if users only monitor certain sub-chains.
This should provide massive scalability with only small drawbacks in security.
\end{abstract}
  

\section{Introduction}

The plasma system \cite{plasma} defines a structure of interconnected blockchains
arranged in a tree structure that promises scalable smart contracts. One of the key
ideas there is that each of the blockchains regularly store their current block hash
in their parent chain so that users can challenge
potentially invalid child state transitions in the parent chain.
This model is secure not because of a difficult proof of work (the chains would
use proof of stake or even a fixed validator set), but rather because
users watch chains they have a stake in and thus will challenge invalid
state transitions, potentially escalating as far up as the trusted main chain.

Here, the scalabality does not come from the fact that blockchains are relieved
from their load by creating a big number of smaller chains and moving the transactions there.
Scalability is only
achieved once a user does not have to verify every single transaction that is
sent to the system. If, for example, a user only cares about a single smart
contract that resides in a single chain that is a leaf of the system,
it is sufficient for the user to verify this leaf and all nodes on the 
path to the root chain. If a transaction is commited by means of block hashes
all the way up to the root chain and there is no invalid state transition in
the chains on the way up to the root, the user can be reasonably sure that
the transaction cannot be declared invalid by other users, essentially
halting at the last ``known good'' state.

This system still does not solve the scalability problem: As long as
the smart contract only ``lives'' inside a single blockchain, it can
merely process a limited amount of transactions. While this might be enough
for some use-cases, a token contract can easily reach this limit.

The system would scale, if the token contract exists on all of the blockchains
and it is possible to move tokens up and down the tree. Users would have accounts
in only one or perhaps some of the chains and watch the paths to the root
from chose chains. In such a simple model, an attacker could just select a
chain that is mostly unused, take it over, create an invalid state transition
that creates tokens out of thin air and then move these tokens up to the root.
If nobody is watching the attacked chain (or the attacker can turn off their
computers or censor their transactions), the attacker is safe as soon as he
or she is able to move the tokens far enough up.

Creating tokens out of thin air is a violation of the assumed invariants
of a token contract. Unfortunately, in contrast to the other invariant
``nobody but me can spend my tokens'', this is a global invariant.
Luckily, this global invariant can be split in a way, such that only enforcing
it locally will still enforce it globally, or at least in the root chain:

We track the sum of the token balances of each direct child chain in the smart
contract of the parent chain and enforce that changes to these numbers
always have to come with a respective amount of tokens being moved from
the child to the parent or vice-versa. In this situation, an attacker can still
create tokens in unwatched child chains, but he or she can only move as many tokens
out of these flawed chains as is their total balance. For users that are not interested
in these chains, the situation would not change: For them, someone took out
tokens from a pool, but it is not relevant who did it.
In effect, the attacker of course steals tokens from users who that have
accounts in the attacked child chains, but at least the impact of the attack is
confined to chains that are not properly watched. In turn, this means that you
have to constantly watch all chains you have tokens in. But there is also
a solution to this: If you do not
want to move your tokens very often, you can transfer them to a chain further
towards the root. These chains likely have higher transaction fees (which
is not relevant if you just want to park your tokens), but also
provide higher security because they are watched by more people.

In the next sections we will give examples of smart contract systems,
how they can be distributed among a plasma system and will also go deeper
into how the synchronisation between parent and child chains is implemented.

\section{Assumptions}

We start by defining a simplified view of the plasma system.

We abstract away the fraud proof mechanism and blockchains in general.

There is a number of \emph{chains}, each of which is modeled as a
computing node. Some chains are malicious, meaning they can have
invalid state transitions, ignore transactions or withhold information.
The chains are arranged in a tree and have a communication channel
to thair parent and their children, if they have some.
We assume that the root chain is not malicious.

There is a certain number of users. Some users are malicous.
We assume that each user can \emph{watch} a limited number of chains.
Non-malicious users watch some chains and
all chains on the path to the root from these chains.

We also assume that if a user watches a chain,
it cannot have any invalid state transitions for that period of time,
unless the user is also malicious or the parent chain can have
invalid state transitions.

State transitions in individual chains are unrelated and happens asynchronous.
Communications between chains are synchronisation points.

\section{Token Contract}

TODO: I wanted to give a formal abstract treatment first,
but it is probably better to argue alongside the contract.

% We want to have a token contract that is deployed to all of the chains.
% Every account has a separate balance on each of the chains and tokens can be
% moved between the chains along the edges in the tree.
% Individually, all chains fulfill the following invariants:

% \begin{enumerate}
% \item If the balance of an account decreases, there has to be a valid signature by the owner.
% \item The sum of all balances in a chain only changes by the amounts sent to and
% received from parents and children.
% \end{enumerate}

% Of course, these invariants can be violated by malicious chains, but we should still have
% the following property:

% \begin{property}
% If a user watches all chains where he or she owns tokens and only accepts to receive
% tokens on chains of which he or she has validated the full history, then the user
% can always move the tokens to the root chain.
% \end{property}
% \begin{proof}

% \end{proof}

The following smart contract is replicated on all chains in
the system. The smart contract language used follows
the syntax and semantic of Solidity~\cite{solidity}, but it has
one additional feature: Functions can be marked ``edge''.
Such functions
are executed as part of a transaction sent to an ``edge''
of the tree instead of a sigle chain. Parts of the code of these functions
are executed on the relative parent and other parts on
the relative child in sequence. The actual compiled smart contract
will use logs and Merkle proofs for synchronisation.
Inside an ``edge'' functions, the identifier
``child'' is an integer (0 or 1 for two children) identifying the
child relative to the parent.

In the event where a child chain is declared faulty, the child
parts can also be executed in the parent chain or recursively
in any chain on the way up to the root (updating the
root hash of the child chain stored in the parent chain and requiring
relevant Merkle proofs for execution). A child that does not react to
edge transactions after a certain time is declared faulty.

\begin{lstlisting}
contract Token {
  mapping(address => uint) balance;
  uint[2] childBalance;

  // Regular transfer between two accounts in the same chain.
  function transfer(address recipient, uint amount) {
    require(balance[msg.sender] >= amount);
    balance[msg.sender] -= amount;
    balance[recipient] += amount;
  }

  // Moves tokens from a certain account in the parent to the
  // same account in the child chain.
  edge function transferToChild(amount) {
      parent {
        require(balance[msg.sender] >= amount);
        balance[msg.sender] -= amount;
        childBalance[child] += amount;
      }
      // After successful completion of the parent part,
      // the child part can be executed in the child.
      child {
        balance[msg.sender] += amount;
      }
  }

  edge function transferToParent(amount) {
    child {
      require(balance[msg.sender] >= amount);
      balance[msg.sender] -= amount;
    }
    parent {
      require(childBalance[child] >= amount);
      childBalance[child] -= amount;
      balance[msg.sender] += amount;
    }
  }
}
\end{lstlisting}

\begin{property}
If a non-malicious user watches all chains where he or she owns tokens and only accepts to receive
tokens on chains where he or she has either validated the full history of the chain
(including the path to the root) or
checked that sum of all balances (including child balances) matches the respective amount in the parent chain
(including the path to the root), then the user
can always move these tokens to the root chain.
\end{property}
\begin{proof}
First, let us observe that if such a user owns accepted tokens in a chain, the user
watches this chain and its parents and thus the code will executed exactly as stated.

Furthermore, note that in each situation where a balance of a user (as opposed
to a child chain) is decreased, the address ``msg.sender'' is used and thus
this only happens if the user intends to do so.

If a child is ignoring edge transactions, it can be declared faulty.
Since the user has been watching the chain, the user can provide Merkle
proofs to move his or her tokens to the parent by executing the edge
transaction only in the parent, potentially recursively until a non-faulty chain is reached.

If the child does not ignore edge transactions and is not declared faulty,
all contracts on the path to the root execute as written.
Then, the only reason the tokens cannot be moved to the root
is that some call to ``transferToParent'' fails. Since the same value of
``amount'' is subtracted in the child chain and added in the parent chain,
the reason for such a failure has to be a failed ``require'' call.

A) The call fails in the child part. The call can only fail in the child if the
user wanted to send more tokens that his or her balance is, but the user
would send exactly ``balance[user]'', so this cannot happen.

B) The call fails in the parent part. In this situation, we have to have
``childBalance[child] $<$ amount''. Remember that the user either verified
the full history of the chain or checked that the balances in the
child sum up to ``childBalance[child]'' in the parent. At least at synchronisation
points, this property is an invariant also for all other functions.
Since ``child.balance[msg.sender] == amount'' at the beginning of the
function call, we have that ``childBalance[child] $\ge$ child.balance[msg.sender]''
and thus a contradiction.

Since both situations are impossible, the user can move the tokens
to the root chain.
\end{proof}

\section{Second Example}

\section{Details of Edge Synchronisation}

Transactions always contain a unique ID of the chain they are sent to
to avoid replays in other chains.


\bibliographystyle{alpha}
\bibliography{bibliography}


\end{document}
