\documentclass[11pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{latexsym,amsbsy,amssymb,amsmath,amsfonts}
\usepackage{amsthm}
\usepackage{epsfig}
\usepackage{euscript}
\usepackage{fullpage}
\usepackage{multirow,multicol,booktabs}
\usepackage[bf]{caption}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{hyperref}
\hypersetup{
 pdfauthor={Christian Reitwießner},
 pdftitle={Casper Slashing Conditions}
  unicode,
  breaklinks,
  colorlinks=false,
  pdfborder={0 0 0}
}

\date{}

\renewcommand{\captionfont}{\small}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\hfuzz=0mm
\tolerance=10000
\hbadness=1000


\setlength{\parindent}{0mm}
\setlength{\parskip}{2ex plus0.5ex minus0.5ex}


\newcommand{\Sum}{\sum\limits}
\newcommand{\Prod}{\prod\limits}


\newtheorem{dummytheorem}{Dummy-Theorem}[section]
\newtheorem{definition}[dummytheorem]{Definition}
\newtheorem{lemma}[dummytheorem]{Lemma}
\newtheorem{theorem}[dummytheorem]{Theorem}
\newtheorem{proposition}[dummytheorem]{Proposition}
\newtheorem{property}[dummytheorem]{Property}
\newtheorem{corollary}[dummytheorem]{Corollary}
\newtheorem{example}[dummytheorem]{Example}
\newtheorem{remark}[dummytheorem]{Remark}
\newtheorem{fact}[dummytheorem]{Fact}
\newtheorem{claim}[dummytheorem]{Claim}
\newtheorem{subclaim}{Subclaim}[dummytheorem]
\newtheorem{conjecture}[dummytheorem]{Conjecture}

\newcommand{\uint}{\mathbb{N}}
\newcommand{\rational}{\mathbb{Q}}

\newcommand{\oli}[1]{\overline{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\newcommand{\eps}{\varepsilon}
\newcommand{\card}[1]{\##1}
\newcommand{\length}[1]{\mathrm{length}({#1})}

\newcommand{\pn}[1]{\textnormal{#1}}

\newcommand{\norm}[1]{||{#1}||}
\newcommand{\Norm}[1]{\left|\left|{#1}\right|\right|}
\newcommand{\eqdef}{:=}

\newcommand{\validators}{\mathcal{V}}
\newcommand{\messages}{\mathcal{M}}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{document}
\selectlanguage{english}


\title{Casper Slashing Conditions}

\author{Christian Reitwießner \\
{\tt chris@ethereum.org}}


\maketitle

\subsection*{Basic Definitions}

Let $\validators$ be the (finite) set of \emph{validators}, where each
$v \in \validators$ has a positive deposit $\mathrm{w}(v)$ ("weight"). We assume neither those
deposits nor the validator set itself to change.
Let $\mathcal{H}$ be the (finite) set of \emph{hash values}, usually
the set of bitstrings of 256 bits. Finally, $\uint \eqdef \{0, 1, \dots\}$
is the set of natural numbers.

Since hashes in $\mathcal{H}$ correspond to blocks in a blockchain, we assume every hash
has a \emph{parent} $\mathrm{parent}(H)$.

\subsection*{Valid Messages}

Every validator can broadcast messages, which can be either \emph{prepare} messages from
the set
\[
  \mathcal{P} \eqdef \{(v, H, n, n_0) \mid
  v \in \validators, H \in \mathcal{H}, n,n_0 \in \uint,
  n > n_0
  \}
\]
or \emph{commit} messages
\[
  \mathcal{C} \eqdef \{(v, H, n) \mid
  v \in \validators, H \in \mathcal{H}, n \in \uint, n > 0\}.
\]

Signatures ensure that only the validator $v$ can send messages $(v, H, n, n_0)$
and $(v, H, n)$.

\subsection*{Slashing Conditions}

Let $\messages \subseteq \mathcal{P} \cup \mathcal{C}$ be the set of messages
visible to the Casper contract at a certain point in time. Depending on
this set, the contract will slash the deposit of validators. To ease notation,
let us define some notions:

The \emph{prepare ratio} of a hash $H \in \mathcal{H}$ at a view $n \in \uint$ depending
on the view $n_0 \in \uint$ is
\[
\mathrm{prepratio}_{\messages}(H, n, n_0) =
\frac{\sum\{w(v) \mid (v, H, n, n_0) \in \messages \}}{\sum_{v \in \validators}\mathrm{w}(v)}
\]
and the prepare ratio of $H$ at the view $n$ is
\[
\mathrm{prepratio}_{\messages}(H, n) = \max_{n_0 \in \uint} \mathrm{prepratio}_\messages(H, n, n_0)
\]

The Casper contract slashes the deposit of a validator $v_0$ if any of the following conditions
are met:

\begin{enumerate}
\item $(v_0, H, n) \in \messages$ for some $H \in \mathcal{H}$, $n \in \uint$,
and $\mathrm{prepratio}_\messages(H, n) < \frac{2}{3}$.\\
-- A hash was commited that was not properly prepared.
\item $(v_0, H, n, n_0) \in \messages$ for some $H \in \mathcal{H}$, $n \in \uint$,
$n_0 > 0$ and $\mathrm{prepratio}_\messages(\mathrm{parent}^{n-n_0}(H), n_0) < \frac{2}{3}$.\\
-- A hash was prepared based on an ancestor that was not properly prepared.
\item $(v_0, H, n), (v_0, H', n', n_0') \in \messages$ for some $H,H' \in \mathcal{H}$, $n, n', n_0' \in \uint$
and $n_0' < n < n'$.\\
-- A hash was prepared ignoring an already committed hash.
\item $(v_0, H, n, n_0), (v_0, H', n, n_0') \in \messages$ for some $H, H' \in \mathcal{H}$,
$n, n_0, n_0' \in \uint$ and $(H, n_0) \neq (H', n_0')$.
-- Two different prepare messages were sent for the same view.
\end{enumerate}

Defined $\mathrm{slashed}_\messages(v_0)$ to be true if and only if at least
one of these conditions are met for $v_0$.

\subsection*{Properties}

\begin{conjecture}[Accountable Safety]
If $(v_1, X, n_1), (v_2, Y, n_2) \in \messages$, $X \neq Y$ and there is no
$k \in \uint$ such that $X = \mathrm{parent}^k(Y)$ or $Y = \mathrm{parent}^k(X)$,
then
\[
  \sum\{\mathrm{w}(v) \mid v \in \mathcal{V}, \mathrm{slashed}_\messages(v)\}
  \ge \frac{1}{3} \sum\{\mathrm{w}(v) \mid v \in \mathcal{V}\}.
\]
\end{conjecture}

\begin{conjecture}[Plausible Liveness]
Let $\messages \subseteq \mathcal{C} \cup \mathcal{P}$ be finite such that
less than a third of the validators are slashed, i.e.
\[
  \sum\{\mathrm{w}(v) \mid v \in \mathcal{V}, \mathrm{slashed}_\messages(v)\} < \frac{1}{3} \sum\{\mathrm{w}(v) \mid v \in \mathcal{V}\}.
\]
Then there is a set of messages $\messages' \supseteq \messages$ and a hash
$H \in \mathcal{H}$ such that
\begin{enumerate}
\item $(v, X, n) \in \messages' \setminus \messages \Rightarrow \neg \mathrm{slashed}_\messages(v)$
(only contains new messages from unslashed validators)
\item there is no $(v, H, n) \in \messages$ ($H$ has not been commited previously)
\item there is some $(v, H, n) \in \messages'$ ($H$ is commited now)
\item for all $v \in \validators$ if $\mathrm{slashed}_{\messages'}(v)$
then $\mathrm{slashed}_{\messages}(v)$ (no newly slashed validator)
\end{enumerate}
\end{conjecture}

\end{document}
